A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-12-19 cusp_j/paper/paper.tex
modified stuffs

Even if these selection schemes are extremely efficient to obtain a qualitative description of the wave function, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. As shown by the seminal works of Hylleraas \cite{Hyl-ZP-29} and further developed byKutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

@@

As the function $u(r_{12};\mu)$ contains a unique parameter, one can try to find the value of $\mfrogg$ such that 
$u(r_{12};\mfrogg)$ qualitatively reproduces the FROGG. 
To do so, we propose a one-point fit such that 
\begin{equation}
 \label{mu_ten_no_1}
u(r_{12}=0;\mfrogg) = f_{\text{TN}}(r_{12}=0), 
\end{equation}
where $ f_{\text{TN}}(x)$ is the FROGG introduced by Ten-No
\begin{equation}
 \label{eq_frogg}
 f_{\text{TN}}(x) = -\sum_{i=1}^6 c_i e^{-\alpha_i x^2 },
\end{equation}
with the coefficients $c_i$ and exponents $\alpha_i$ given in Table 1 of Ref. \onlinecite{TenNo-CPL-00-a}. 
Eq. \eqref{mu_ten_no_1} gives 
\begin{equation}
 \label{mu_ten_no_2}
 \mfrogg = -\frac{1}{2\sqrt{\pi}f_{\text{TN}}(0)},
\end{equation}
which gives $\mfrogg\approx0.86975$. 
We represent on Fig \ref{fig_j_mu} the comparison between $u(r_{12};\mfrogg)$ and the FROGG, and it is clearly apparent  that $u(r_{12};\mfrogg)$ reproduces almost perfectly the FROGG, although it decays slightly faster at large $r_{12}$. Therefore, one can see the FROGG as a special case of $u(r_{12};\mu)$  with $\mu = \mfrogg$.

\subsection{The general analytical form of  $\tilde{H}[\mu]$ and some of its properties}
\label{sec:ht_general}
\subsubsection{General form of $\tilde{H}[\mu]$ }
Having established the analytical expression of $u(r_{12};\mu)$, one can derive the expression of the associated transcorrelated Hamiltonian for a general $N$-electron system.  
According to Eq. (2) of Ref. \onlinecite{CohLuoGutDobTewAla-JCP-19}, the similarity transformed Hamiltonian can be written as 
\begin{equation}
 \label{ht_def_g}
 e^{-\hat{\tau}} \hat{H} e^{\hat{\tau}} = H + \big[ H,\hat{\tau} \big] + \frac{1}{2}\bigg[ \big[H,\hat{\tau}\big],\hat{\tau}\bigg]
\end{equation}
where $\hat{\tau} = \sum_{i<j}u(\br{}_i,\br{}_j)$ and $\hat{H} = \sum_i -\frac{1}{2} \nabla^2_i + v(\br{}_i) + \sum_{i<j} \frac{1}{r_{ij}}$. 
This leads to the following similarity transformed Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H} & = H - \sum_i \bigg( \frac{1}{2} \nabla_i^2 \hat{\tau} + \big(\nabla_i \hat{\tau} \big) + \frac{1}{2} \big(\nabla_i \hat{\tau} \big)^2  \bigg) \\
           & = H - \sum_{i<j} \hat{K}(\bri{i},\bri{j}) - \sum_{i<j<k} \hat{L}(\bri{i},\bri{j},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}(\bri{i},\bri{j})$ and $\hat{L}(\bri{i},\bri{j},\bri{k})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}(\bri{1},\bri{2})  = \frac{1}{2} \bigg( &\nabla_1^2 u(\bri{1},\bri{2}) + \nabla_2^2u(\bri{1},\bri{2}) \\
                                               + &\big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 + \big(\nabla_1 u(\bri{1},\bri{2}) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_2 + \nabla_2 u(\bri{1},\bri{2}) \cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}(\bri{1},\bri{2},\bri{3})  = & \nabla_1 u(\bri{1},\bri{2}) \cdot \nabla_1 u(\bri{1},\bri{3}) + \nabla_2 u(\bri{2},\bri{1}) \cdot \nabla_2 u(\bri{2},\bri{3})  \\
                                     + & \nabla_3 u(\bri{3},\bri{1}) \cdot \nabla_3 u(\bri{3},\bri{2}).
 \end{aligned}
\end{equation}
Therefore, to obtain the analytical expression of the $\tilde{H}[\mu]$ for a general $N$-electron system, 
one needs to compute the actual form of the operators $\hat{K}(\bri{1},\bri{2})$ and $\hat{L}(\bri{1},\bri{2},\bri{3})$ when $u(\bri{1},\bri{2}) = u(r_{12},\mu)$.

\subsection{Analytical form for the operators $\hat{K}_\mu(\bri{1},\bri{2}) $ and $\hat{L}_\mu(\bri{1},\bri{2},\bri{3})$ }
To compute the operators $\hat{K}_\mu(\bri{1},\bri{2})$ and $\hat{L}_\mu(\bri{1},\bri{2},\bri{3}) $ one needs to compute first the gradient of $u(r_{12},\mu)$ with respect to $\bri{1}$ which reads 
\begin{equation}
 \label{eq:nabla_u}
 \begin{aligned}
 \nabla_1 u(r_{12},\mu) = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \big( \br{}_1 - \br{}_2 \big).
 \end{aligned}
\end{equation}
Then, the term $\big(\nabla_1 u(r_{12},\mu) \big) ^2 + \big(\nabla_2 u(r_{12},\mu) \big) ^2$ is simply 
\begin{equation}
 \begin{aligned}
 \big(\nabla_1 u(r_{12},\mu) \big) ^2 + \big(\nabla_2 u(r_{12},\mu) \big) ^2 = \frac{\bigg(1 - \text{erf}(\mu r_{12}) \bigg)^2}{2}.
 \end{aligned}
\end{equation}
The non hermitian operators in $\hat{K}_\mu(\bri{1},\bri{2})$ contains 
\begin{equation}
 \begin{aligned}
 \nabla_1 u(r_{12},\mu) \cdot \nabla_1  = &\frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\ 
                                          &\bigg( (x_1 - x_2) \deriv{}{x_1}{} + (y_1 - y_2) \deriv{}{y_1}{} + (z_1 - z_2) \deriv{}{z_1}{}\bigg),
 \end{aligned}
\end{equation}
and as $\nabla_1 u(r_{12},\mu) = - \nabla_2 u(r_{12},\mu)$,   
the total non hermitian operator in $\hat{K}_\mu(\bri{1},\bri{2})$ can be written as 
\begin{equation}
 \begin{aligned}
 \label{def_non_hermit}
& \nabla_1 u(r_{12},\mu) \cdot \nabla_1 + \nabla_2 u(r_{12},\mu) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \\
& \bigg( (x_1 - x_2) \big( \deriv{}{x_1}{} - \deriv{}{x_2}{} \big) +
         (y_1 - y_2) \big( \deriv{}{y_1}{} - \deriv{}{y_2}{} \big)  \\
&  +      (z_1 - z_2) \big( \deriv{}{z_1}{} - \deriv{}{z_2}{} \big)\bigg).
 \end{aligned}
\end{equation}
One can notice that the form of Eq. \eqref{def_non_hermit} provides an explicit form to compute analytical integrals with Gaussian types functions. 
Nevertheless, one can notice that as 
\begin{equation}
 \deriv{}{r_{12}^x}{} = \frac{1}{2} \bigg( \deriv{}{x_1}{} - \deriv{}{x_2}{} \bigg),
\end{equation}
one can write 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla0}
& \nabla_1 u(r_{12},\mu) \cdot \nabla_1 + \nabla_2 u(r_{12},\mu) \cdot \nabla_2 = \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}}.
 \end{aligned}
\end{equation}
Then, introducing the spherical coordinate system for $\br{}_1 - \br{}_2$ in terms of ${\bf e}_u= \frac{\br{}_1 - \br{}_2}{r_{12}}$ and the angle $\theta$ and $\phi$, one can write 
\begin{equation}
 \nabla_{\br{}_{12}} = \deriv{}{r_{12}}{} {\bf e}_u + \frac{1}{r_{12}} \deriv{}{\theta}{} {\bf e}_{\theta} + \frac{1}{r_{12} \sin(\theta)} \deriv{}{\phi}{} {\bf e}_\phi,
\end{equation}
and as $\br{}_1 - \br{}_2 = r_{12} {\bf e}_u$ one obtains
\begin{equation}
 \big( \br{}_1 - \br{}_2 \big) \cdot \nabla_{\br{}_{12}} = r_{12} \deriv{}{r_{12}}{}.
\end{equation}
Therefore 
\begin{equation}
 \begin{aligned}
 \label{eq:nabla_i_nabla1}
 \nabla_1 u(r_{12},\mu) \cdot \nabla_1 + \nabla_2 u(r_{12},\mu) \cdot \nabla_2 = \bigg( 1 - \text{erf}(\mu r_{12})\bigg) \deriv{}{r_{12}}{},
 \end{aligned}
\end{equation}
which is a more compact expression than Eq.\eqref{def_non_hermit}

The computation of the Laplacian reads 
\begin{equation}
 \begin{aligned}
 \label{eq:d2_x1_2}
 &\nabla_1^2 u(r_{12},\mu) + \nabla_2^2u(r_{12},\mu)
 & = 2 \times \bigg( \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2}  \bigg).
 \end{aligned}
\end{equation}

Summing all terms to recombine $\hat{K}_\mu(\bri{1},\bri{2})$ reads 
\begin{equation}
 \label{eq:k_final}
  \begin{aligned}
   \hat{K}_\mu(\bri{1},\bri{2}) = & \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} + \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4} \\
& - \bigg( \text{erf}(\mu r_{12}) - 1\bigg) \deriv{}{r_{12}}{}.
  \end{aligned}
\end{equation}
Similarly, the explicit analytical form of $\hat{L}_\mu(\bri{1},\bri{2},\bri{3}) $ reads 
\begin{equation}
 \label{eq:l_final}
 \begin{aligned}
 \hat{L}_\mu(\bri{1},\bri{2},\bri{3}) = & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{12} \cdot \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{13} \\
                                      + & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{21} \cdot \frac{1 - \text{erf}(\mu r_{23})}{2 r_{23}} \bri{23} \\
                                      + & \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{31} \cdot \frac{1 - \text{erf}(\mu r_{32})}{2 r_{32}} \bri{32}.
 \end{aligned}                    
\end{equation}

\subsection{Analytical form of $\tilde{H}[\mu]$}
Having established the explicit form of $\hat{K}\mu(\bri{1},\bri{2})$ and $\hat{L}_\mu(\bri{1},\bri{2},\bri{3})$, one can now 
look at the explicit analytical form of the transcorrelated Hamiltonian $\tilde{H}[\mu]$ obtained with $u(r_{12};\mu)$,  
which, by injecting the form of Eqs. \eqref{eq:k_final} and \eqref{eq:l_final} in Eq.\eqref{ht_def_gu}, gives
\begin{equation}
  \begin{aligned}
 \label{eq:exp_ht_mu}
   \tilde{H}[\mu] = & h_c + \sum_{i<j} \bigg( \tilde{\mathcal{W}}_{ee}(r_{ij},\mu) + \tilde{t}[\mu] \bigg) - \sum_{i<j<k} \hat{L}_\mu(\bri{1},\bri{2},\bri{3})
  \end{aligned}
\end{equation}
where $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ is the scalar effective two-body potential of  $\tilde{H}[\mu]$
\begin{equation}
 \begin{aligned}
 \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)  =  \frac{\text{erf}(\mu r_{ij})}{r_{ij}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{ij} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{ij}) \bigg)^2}{4}, 
 \end{aligned}
\end{equation}
and $\tilde{t}[\mu]$ is a non hermitian two-body differential operator 
\begin{equation}
 \begin{aligned}
 \tilde{t}[\mu] =  \bigg( \text{erf}(\mu r_{ij}) - 1\bigg) \deriv{}{r_{ij}}{}.
 \end{aligned}
\end{equation}

@@

As the Jastrow factor $u(r_{12},\mu)$ for $\mu=\mfrogg$ is very close to the FROGG, we can also compare the scalar effective potential $\tilde{\mathcal{W}}_{ee}(r_{12},\mfrogg)$ with that obtained with the FROGG which is given by
\begin{equation}
 \label{eq:_wee_frogg}
 \tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})= \frac{1}{r_{12}}-\nabla_1^2 f_{\text{TN}}(r_{12}) - \big(\nabla_1 f_{\text{TN}}(r_{12}) \big) ^2. 
\end{equation}
As apparent from Fig. \ref{fig_wee_frogg} where $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ and $\tilde{\mathcal{W}}_{ee}(r_{12},\mfrogg)$ are represented, while the latter is perfectly smooth for all $r_{12}$, the former strongly oscillates near $r_{12}=0$ between positive and negative values and is not bounded for $r_{12}=0$. 
Such behaviour of $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ near $r_{12}=0$ comes from the fact that,
as it is made of gaussian functions, $f_{\text{TN}}(r_{12})$ is cusp-less \textit{i.e.} 
\begin{equation}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}f_{\text{TN}}(r_{12}) = 0, 
\end{equation}
and as a necessary condition to obtain a non divergent scalar potential is 
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12}) = \frac{1}{2}, 
\end{equation}
(see Eqs. \eqref{eq:cusp_phi_0}  and \eqref{eq:cusp_phi_1} in Appendix \ref{sec:cusp}), $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ is necessary not bounded at $r_{12}=0$.  
%Because the condition of Eq. \eqref{eq:cusp_phi_0} is fulfilled by $u(r_{12},\mu)$ for all $\mu < \infty$, 
%the effective potential obtained by the similarity transformation of $H$ by $u(r_{12},\mu)$ is necessary non divergent.

@@

\subsection{Limit cases for $\tilde{H}[\mu]$ with $\mu$}
\label{sec:h_mu_lim}
Regarding now the variation of $\tilde{H}[\mu]$, as 
\begin{equation}
 \label{eq:lim_mu_0}
 \lim_{\mu  \rightarrow \infty }u(\mu,r_{12}) = 0, 
\end{equation}
one recovers the usual Hamiltonian in the ${\mu  \rightarrow \infty }$ limit:
\begin{equation}
 \label{eq:lim_mu_1}
 \lim_{\mu \rightarrow \infty} \tilde{H}[\mu] = H,
\end{equation}
although it is only in the sense of distributions (see Appendix \ref{sec:large_mu_lim} for a more detailed explanation).

@@

Also, it is striking to see that, in a given basis set, the ground state eigenvalue of the TC Hamiltonian obtained by the FROGG and that obtained by $\tilde{H}[\mu]$ with $\mu=\mfrogg$ coincide with less than 1 mH. 
The latter results confirms that as $u(r_{12},\mfrogg) \approx f_{\text{TN}}(r_{12})$ (see Fig \ref{fig_j_mu} and Eqs. \eqref{mu_ten_no_1} and \eqref{mu_ten_no_2}), the TC Hamiltonians obtained with the two Jastrow factor are very similar, as illustrated for instance by the effective potential in Fig. \ref{fig_wee_frogg}.

@@

From the study of the ground state of the helium atom, one can conclude that i) the use of a simple Jastrow factor such as $u(r_{12};\mu)$ within the transcorrelated framework can strongly improve the speed of convergence of the total energy of the helium atom with a quite wide range of values of $\mu$ ranging from $\mu=0.2$ to $\mu=1.6$, 
ii) the results obtained with $\mu=\mfrogg$ reproduce essentially that obtained with the FROGG, 
iii) for $\mu \in[ 0.35,0.7]$ the results obtained are always better than that obtained with the FROGG. 
\begin{figure*}
 \label{fig_conv_e_mu}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_small_mu.pdf}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_large_mu.pdf}\\
        \caption{Convergence of the ground state eigenvalue of $\tilde{H}[\mu]$ with respect to the AVXZ basis set series (X=D,T,Q,5) for different values of $\mu$. 
We also compare with the results of the TC obtained with the $u(r_{12}) = f_{\text{TN}}(r_{12})$ (see Eq. \eqref{eq_frogg}), referred here after as FROGG. }
\end{figure*}

>>>>

Even if these selection schemes are extremely efficient to obtain a qualitative description of the wave function, a quantitative description is often out of reach as it implies the use of very large one-electron basis sets which are needed to recover the slowly converging correlation effects near the electron-electron coalescence point. 
These short-range correlation effects are mainly due to the diverging character of the coulomb potential, which induces a non analytical behaviour of the wave function at small inter-electronic distance ($r_{12}\approx 0$): the famous electronic cusp originally derived by Kato\cite{Kat-CPAM-57}. For a general and clear derivation of the cusp conditions for Coulombic systems up to second order, see Ref.\onlinecite{Tew-JCP-08}. As shown by the seminal works of Hylleraas\cite{Hyl-ZP-29} and further developed byKutzelnigg and coworkers, \cite{Kut-TCA-85,KutKlo-JCP-91, NogKut-JCP-94} the slow convergence of the correlation energy with respect to the quality of the basis set is mainly due to the impossibility to represent the cusp in a finite one-electron basis set. 
As the cusp conditions originate from the divergence of the Coulomb potential at $r_{12}=0$, an alternative approach would be to develop theories dealing with a smooth potential. Such a path was followed by three distinct branches: 
the explicitly correlated methods (F12), transcorrelated methods (TC) and the range-separated density functional theory (RS-DFT).

@@

As the function $u(r_{12};\mu)$ contains a unique parameter, one can try to find the value of $\mfrogg$ such that 
$u(r_{12};\mfrogg)$ qualitatively reproduces the FROGG. 
To do so, we propose a one-point fit such that 
\begin{equation}
 \label{mu_ten_no_1}
u(r_{12}=0;\mfrogg) = u^{\text{TN}}(r_{12}=0), 
\end{equation}
where $ u^{\text{TN}}(x)$ is the FROGG introduced by Ten-No
\begin{equation}
 \label{eq_frogg}
 u^{\text{TN}}(x) = -\sum_{i=1}^6 c_i e^{-\alpha_i x^2 },
\end{equation}
with the coefficients $c_i$ and exponents $\alpha_i$ given in Table 1 of Ref. \onlinecite{TenNo-CPL-00-a}. 
Eq. \eqref{mu_ten_no_1} gives 
\begin{equation}
 \label{mu_ten_no_2}
 \mfrogg = -\frac{1}{2\sqrt{\pi}u^{\text{TN}}(0)},
\end{equation}
which gives $\mfrogg\approx0.86975$. 
We represent on Fig \ref{fig_j_mu} the comparison between $u(r_{12};\mfrogg)$ and the FROGG, and it is clearly apparent  that $u(r_{12};\mfrogg)$ reproduces almost perfectly the FROGG, although it decays slightly faster at large $r_{12}$. Therefore, one can see the FROGG as a special case of $u(r_{12};\mu)$  with $\mu = \mfrogg$.

\subsection{The general analytical form of  $\tilde{H}[\mu]$ and some of its properties}
\label{sec:ht_general}
\subsubsection{General analytical form of $\tilde{H}[\mu]$ for a $N$-electron system}
Having established the analytical expression of $u(r_{12};\mu)$, one can derive the expression of the associated transcorrelated Hamiltonian for a general $N$-electron system.  
Using the general form of the transcorrelated Hamiltonian of Eq. (2) in Ref. \onlinecite{CohLuoGutDobTewAla-JCP-19} and using $u(r_{12},\mu)$ as Jastrow factor gives 
\begin{equation}
 \begin{aligned}
 \label{ht_def_g}
 \tilde{H}[\mu] &\equiv e^{-\hat{\tau}_\mu} \hat{H} e^{\hat{\tau}_\mu} \\ 
                &= H + \big[ H,\hat{\tau}_\mu \big] + \frac{1}{2}\bigg[ \big[H,\hat{\tau}_\mu\big],\hat{\tau}_\mu\bigg],
 \end{aligned}
\end{equation}
with $\hat{\tau}_\mu = \sum_{i<j}u(r_{ij},\mu)$ and $\hat{H} = \sum_i -\frac{1}{2} \nabla^2_i + v(\br{}_i) + \sum_{i<j} \frac{1}{r_{ij}}$. 
Eq. \eqref{ht_def_g} leads to the following transcorrelated Hamiltonian 
\begin{equation}
 \begin{aligned}
 \label{ht_def_gu}
 \tilde{H}[\mu] & = H - \sum_{i<j} \hat{K}_\mu(\bri{i},\bri{j}) - \sum_{i<j<k} \hat{L}_\mu(\bri{i},\bri{j},\bri{k}),
 \end{aligned}
\end{equation}
where the effective two- and three-body operators $\hat{K}_\mu(\bri{1},\bri{2})$ and $\hat{L}_\mu(\bri{1},\bri{2},\bri{3})$ are defined as
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{K}_\mu(\bri{1},\bri{2})  = \frac{1}{2} \bigg( &\Delta_1 u(r_{12},\mu) + \Delta_2 u(r_{12},\mu) \\
                                               + &\big(\nabla_1 u(r_{12},\mu) \big) ^2 + \big(\nabla_1 u(r_{12},\mu) \big) ^2 \bigg) \\
                                               + &\nabla_1 u(r_{12},\mu) \cdot \nabla_2 + \nabla_2 u(r_{12},\mu)\cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}_\mu(\bri{1},\bri{2},\bri{3})  = & \nabla_1 u(r_{12},\mu) \cdot \nabla_1 u(r_{13},\mu) + \nabla_2 u(r_{12},\mu) \cdot \nabla_2 u(r_{12},\mu)  \\
                                     + & \nabla_3 u(r_{12},\mu) \cdot \nabla_3 u(r_{12},\mu).
 \end{aligned}
\end{equation}

As shown in the Appendix \ref{k_l_appendix}, the operator $\hat{K}_\mu(\bri{1},\bri{2})$ has a straightforward analytical expression which reads 
\begin{equation}
 \label{eq:k_final}
  \begin{aligned}
   \hat{K}_\mu(\bri{1},\bri{2}) = & \frac{1 - \text{erf}(\mu r_{12})}{r_{12}} - \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{12} \big)^2} + \frac{\bigg(1 -     \text{erf}(\mu r_{12}) \bigg)^2}{4} \\
& - \bigg( \text{erf}(\mu r_{12}) - 1\bigg) \deriv{}{r_{12}}{}.
  \end{aligned}
\end{equation}
and similarly that of $\hat{L}_\mu(\bri{1},\bri{2},\bri{3}) $ reads 
\begin{equation}
 \label{eq:l_final}
 \begin{aligned}
 \hat{L}_\mu(\bri{1},\bri{2},\bri{3}) = & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{12} \cdot \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{13} \\
                                      + & \frac{1 - \text{erf}(\mu r_{12})}{2 r_{12}} \bri{21} \cdot \frac{1 - \text{erf}(\mu r_{23})}{2 r_{23}} \bri{23} \\
                                      + & \frac{1 - \text{erf}(\mu r_{13})}{2 r_{13}} \bri{31} \cdot \frac{1 - \text{erf}(\mu r_{32})}{2 r_{32}} \bri{32}.
 \end{aligned}                    
\end{equation}

By injecting the form of Eqs. \eqref{eq:k_final} and \eqref{eq:l_final} in Eq.\eqref{ht_def_gu}, one obtains 
\begin{equation}
  \begin{aligned}
 \label{eq:exp_ht_mu}
   \tilde{H}[\mu] = & h_c + \sum_{i<j} \bigg( \tilde{\mathcal{W}}_{ee}(r_{ij},\mu) + \tilde{t}[\mu] \bigg) - \sum_{i<j<k} \hat{L}_\mu(\bri{1},\bri{2},\bri{3})
  \end{aligned}
\end{equation}
where $ \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)$ is the scalar effective two-body potential of  $\tilde{H}[\mu]$
\begin{equation}
 \begin{aligned}
 \tilde{\mathcal{W}}_{ee}(r_{ij},\mu)  =  \frac{\text{erf}(\mu r_{ij})}{r_{ij}} + \frac{\mu}{\sqrt{\pi}} e^{-\big(\mu r_{ij} \big)^2} - \frac{\bigg(1 -     \text{erf}(\mu r_{ij}) \bigg)^2}{4}, 
 \end{aligned}
\end{equation}
and $\tilde{t}[\mu]$ is a non hermitian two-body differential operator 
\begin{equation}
 \begin{aligned}
 \tilde{t}[\mu] =  \bigg( \text{erf}(\mu r_{ij}) - 1\bigg) \deriv{}{r_{ij}}{}.
 \end{aligned}
\end{equation}

@@

\subsubsection{A special case of $\tilde{H}[\mu]$ which mimics the FROGG}
As the Jastrow factor $u(r_{12},\mu)$ for $\mu=\mfrogg$ is very close to the FROGG, we can also compare the scalar effective potential $\tilde{\mathcal{W}}_{ee}(r_{12},\mfrogg)$ with that obtained with the transcorrelated Hamiltonian obtained with the FROGG. 
The latter is simply defined as 
\begin{equation}
 \label{eq:tn_htilde}
 \tilde{H}_{\text{TN}} = e^{-\hat{\tau}_{\text{TN}}}H e^{\hat{\tau}_{\text{TN}}},
\end{equation}
with $\hat{\tau}_{\text{TN}} = \sum_{i<j} u^\text{TN}(r_{ij})$. 
This leads to the following transcorrelated Hamiltonian 
\begin{equation}
 \label{eq:tn_htilde_2}
  \tilde{H}_{\text{TN}} = H + \sum_{i<j} \bigg( \tilde{\mathcal{W}}^\text{TN}_{ee}(r_{ij}) + \tilde{t}^\text{TN} \bigg) - \sum_{i<j<k} \hat{L}^\text{TN}(\bri{1},\bri{2},\bri{3})
\end{equation}
where 
\begin{equation}
 \label{eq:_wee_frogg}
 \tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})= \frac{1}{r_{12}}-\Delta_1 u^{\text{TN}}(r_{12}) - \big(\nabla_1 u^{\text{TN}}(r_{12}) \big) ^2, 
\end{equation}
\begin{equation}
 \begin{aligned}
 \tilde{t}^\text{TN} = &\nabla_1 u^\text{TN}(r_{12}) \cdot \nabla_2 + \nabla_2 u^\text{TN}(r_{12})\cdot \nabla_1,
 \end{aligned}
\end{equation}
\begin{equation}
 \begin{aligned}
 \label{def_k}
  \hat{L}_\mu(\bri{1},\bri{2},\bri{3})  = & \nabla_1 u^\text{TN}(r_{12}) \cdot \nabla_1 u^\text{TN}(r_{12}) + \nabla_2 u^\text{TN}(r_{12}) \cdot \nabla_2 u^\text{TN}(r_{12})  \\
                                     + & \nabla_3 u^\text{TN}(r_{12}) \cdot \nabla_3 u^\text{TN}(r_{12}).
 \end{aligned}
\end{equation}
As apparent from Fig. \ref{fig_wee_frogg} where $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ and $\tilde{\mathcal{W}}_{ee}(r_{12},\mfrogg)$ are represented, while the latter is perfectly smooth for all $r_{12}$, the former strongly oscillates near $r_{12}=0$ between positive and negative values and is not bounded for $r_{12}=0$. 
Such behaviour of $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ near $r_{12}=0$ comes from the fact that,
as it is made of gaussian functions, $u^{\text{TN}}(r_{12})$ is cusp-less \textit{i.e.} 
\begin{equation}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u^{\text{TN}}(r_{12}) = 0, 
\end{equation}
and as a necessary condition to obtain a non divergent scalar potential is 
\begin{equation}
 \label{eq:cusp_phi_0}
 \lim_{r_{12}\rightarrow 0} \deriv{}{r_{12}}{}u(r_{12}) = \frac{1}{2}, 
\end{equation}
(see Eqs. \eqref{eq:cusp_phi_0}  and \eqref{eq:cusp_phi_1} in Appendix \ref{sec:cusp}), $\tilde{\mathcal{W}}_{ee}^\text{TN}(r_{12})$ is necessary not bounded at $r_{12}=0$.  
%Because the condition of Eq. \eqref{eq:cusp_phi_0} is fulfilled by $u(r_{12},\mu)$ for all $\mu < \infty$, 
%the effective potential obtained by the similarity transformation of $H$ by $u(r_{12},\mu)$ is necessary non divergent.

@@

\subsubsection{Limit cases for $\tilde{H}[\mu]$ with $\mu$}
\label{sec:h_mu_lim}
Regarding now the variation of $\tilde{H}[\mu]$, as 
\begin{equation}
 \label{eq:lim_mu_0}
 \lim_{\mu  \rightarrow \infty }u(\mu,r_{12}) = 0, 
\end{equation}
one recovers the usual Hamiltonian in the ${\mu  \rightarrow \infty }$ limit:
\begin{equation}
 \label{eq:lim_mu_1}
 \lim_{\mu \rightarrow \infty} \tilde{H}[\mu] = H,
\end{equation}
although it is only in the sense of distributions (see Appendix \ref{sec:large_mu_lim} for a more detailed explanation).

@@

Also, it is striking to see that, in a given basis set, the ground state eigenvalue of the TC Hamiltonian obtained by the FROGG and that obtained by $\tilde{H}[\mu]$ with $\mu=\mfrogg$ coincide with less than 1 mH. 
The latter results confirms that as $u(r_{12},\mfrogg) \approx u^{\text{TN}}(r_{12})$ (see Fig \ref{fig_j_mu} and Eqs. \eqref{mu_ten_no_1} and \eqref{mu_ten_no_2}), the TC Hamiltonians obtained with the two Jastrow factor are very similar, as illustrated for instance by the effective potential in Fig. \ref{fig_wee_frogg}.

@@

From the study of the ground state of the helium atom, one can conclude that i) the use of a simple Jastrow factor such as $u(r_{12};\mu)$ within the transcorrelated framework can strongly improve the speed of convergence of the total energy of the helium atom with a quite wide range of values of $\mu$ ranging from $\mu=0.2$ to $\mu=1.6$, 
ii) the results obtained with $\mu=\mfrogg$ reproduce essentially that obtained with the FROGG, 
iii) for $\mu \in[ 0.35,0.7]$ the results obtained are always better than that obtained with the FROGG. 
\begin{figure*}
 \label{fig_conv_e_mu}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_small_mu.pdf}
        \includegraphics[width=0.45\linewidth]{plots/He/Energy/He_E_conv_basis_large_mu.pdf}\\
        \caption{Convergence of the ground state eigenvalue of $\tilde{H}[\mu]$ with respect to the AVXZ basis set series (X=D,T,Q,5) for different values of $\mu$. 
We also compare with the results of the TC obtained with the $u(r_{12}) = u^{\text{TN}}(r_{12})$ (see Eq. \eqref{eq_frogg}), referred here after as FROGG. }
\end{figure*}
